Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

### 2023-10-12 manuscript2.RMD
updates

Nutritional supplements such as protein and creatine is often used to optimise RT adaptations [@cermak_protein_2012;@lanhers_creatine_2015;@lanhers_creatine_2017;@morton_systematic_2018]. However, it remains equivocal if other nutritional adjuvants such as glucose can increase the efficacy of RT [@tezze_anabolic_2023]. This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major energy supplier to cells via adenosine triphosphate (ATP) synthesis [@mul_exercise_2015;@sebata_control_2018]. Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019]. In addition, insulin may have anabolic effects with elevated amino acids, and has a clear role in reduction of muscle protein breakdown independent of amino acid availability [@abdulla_role_2016]. Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the potential additive effect of glucose and RT on ribosome biogenesis.

@@

To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle, making our hypotheses exploratory and comparisons few. *In vitro* studies on mice glomerular epithelial cells [@mariappan_ribosomal_2011] and human breast cancer cells [@tanaka_mild_2015] have observed glucose-induced UBF-dependent augmentation of ribosome biogenesis [@mariappan_ribosomal_2011] and reduced rRNA transcription by glucose starvation [@tanaka_mild_2015]. These findings do imply an important role for glucose in the initiation of rDNA transcription, the rate-limiting step of ribosome biogenesis [@mariappan_ribosomal_2011; @tanaka_mild_2015]. In addition, Hillier et al. found physiological hyperinsulinemia to stimulate p70S6K phosphorylation in human skeletal muscle [@hillier_physiological_2000]. Moreover, robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA can be expected after an acute bout of RT [@figueiredo_impact_2016] or a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022]. Therefore, we hypothesized that if glucose can, directly or indirectly via insulin, stimulate the initiation of rDNA transcription, five sessions of RT with glucose compared to placebo would elicit this effect. Further, glucose ingestion increases energy levels, leading us to hypothesize that high glucose vs. placebo condition would stimulate energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) and Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012]. Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation  [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study showed no signs of such effects of glucose vs. placebo conditions. Resistance training, irrespective of condition, yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observations [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022]. Due to the exploratory nature of our hypotheses, this was hardly a very surprising result. Indeed, previous studies have used cell cultures from yeast [@zhai_human_2012], rodents [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells. Importantly, previous studies investigated high vs. low glucose conditions [@mariappan_ribosomal_2011], or high glucose vs. glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare the high glucose condition to a placebo condition, with a matched daily macronutrient and energy intake. Therefore, the comparison made in the present study was high plasma glucose levels vs. normal plasma glucose levels, to investigate the effect of glucose *per se* and not intracellular energy status. Thus, while glucose ingestion presumably is important for supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis.

In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target UBF and the ribosomal protein S6 (rpS6), as well as the general transcription factor c-Myc, further supported observations from the RNA data as there was no difference between the glucose and placebo conditions. UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 is proposed as a valid and reliable means to measure ribosome biogenesis [@chaillou_hypoxia_2012; @nakada_correlation_2016]. Lastly, c-Myc has previously been described as a potent regulator of ribosome biogenesis, independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016]. Hence, it seems quite reasonable to observe similar changes in these three proteins. The linear relationship found exclusively between UBF content and total RNA levels, and neither c-Myc nor rpS6, supports the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis. While this is not a novel finding, it has only recently been observed in human skeletal muscle following a period RT [@hammarstrom_ribosome_2022]. As such, UBF also seems to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].

>>>>

Nutritional supplements such as protein and creatine is often used to optimise RT adaptations [@cermak_protein_2012;@lanhers_creatine_2015;@lanhers_creatine_2017;@morton_systematic_2018]. However, it remains equivocal if other nutritional adjuvants such as glucose can increase the efficacy of RT [@tezze_anabolic_2023]. This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major energy supplier to cells via adenosine triphosphate (ATP) synthesis [@mul_exercise_2015;@sebata_control_2018]. Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019]. In addition, insulin may have anabolic effects with elevated amino acids, and has a clear role in reduction of muscle protein breakdown independent of amino acid availability [@abdulla_role_2016; @hillier_physiological_2000]. Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the potential additive effect of glucose and RT on ribosome biogenesis.

@@

The observations of the present study, from human skeletal muscle, do not support previous indications from non-human skeletal muscle [@mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012]. Allthough we observed a linear relationship between baseline and trained state UBF and total RNA levels, a UBF-dependent augmentation of ribosome biogenesis did not seem to be induced by glucose ingestion in the present study, as compared to in mice glomerular epithelial cells [@mariappan_ribosomal_2011] and in human breast cancer cells [@tanaka_mild_2015]. Neither do our findings indicate that insulin *per se*, at least in physiological levels, potentiates accumulation of total RNA through p70S6K stimulation as observed with hyperinsulinemia in human skeletal muscle [@hillier_physiological_2000]. Robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA are expected to occur after an acute bout of RT [@figueiredo_impact_2016], as well as after a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022]. Therefore, in the present study, it was expected that eventual benefits of ingesting glucose compared to placebo with RT would be measurable after either 5 training sessions, either due to glucose-induced stimulation of energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) or Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012]. Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation  [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study showed no signs of such effects of glucose vs. placebo conditions. Importantly, previous studies investigated high vs. low glucose conditions [@mariappan_ribosomal_2011], or high glucose vs. glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare the high glucose condition to a placebo condition, with a matched daily macronutrient and energy intake. Therefore, the comparison made in the present study was high plasma glucose levels vs. normal plasma glucose levels, to investigate the effect of glucose *per se* and not intracellular energy status. Thus, while glucose ingestion presumably is important for supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis, during 12 days of heavy-load RT. Further, previous studies have used cell cultures from yeast [@zhai_human_2012], rodents [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human breast cancer cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells. Nevertheless, resistance training irrespective of condition yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observation for heavy-load RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].

In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target UBF and the ribosomal protein S6 (rpS6), as well as the general transcription factor c-Myc, further supported observations from the RNA data as there was no difference between the glucose and placebo conditions. UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 is proposed as a valid and reliable means to measure ribosome biogenesis [@chaillou_hypoxia_2012; @nakada_correlation_2016]. Lastly, c-Myc has previously been described as a potent regulator of ribosome biogenesis, independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016]. Hence, it seems quite reasonable to observe similar changes in these three proteins. The linear relationship found exclusively between UBF content and total RNA levels, and neither c-Myc nor rpS6, supports the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis. While this is not a novel finding, it was only recently observed in human skeletal muscle following a period of RT [@hammarstrom_ribosome_2022]. As such, UBF also seems to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].
